---
title: "Data Analysis Practice"
output: html_document
date: "2024-02-22"
---

> While completing my Masters at Villanova, I will be working in the R.K. Wieder lab on research related to greenhouse gas (GHG) emissions from oil sands development in Alberta, Canada, and their reflection in the surrounding environment of peatland habitat. As of now, in my first semester at Villanova, I don't have data of my own. To create these charts I am using randomized data and data from previous years to showcase my progress with R and practice making figures for my future research!

## Randomized Data Set: 

```{r}
numeric_data <- data.frame(
  Year = 1924:2024,
  Temperature = runif(101, min = 40, max = 80) # Generates 101 random numbers between 40 and 80
)

write.csv(numeric_data, file = "numeric_data.csv", row.names = FALSE)

# Filter data for last 10 years
last_10_years <- subset(numeric_data, Year >= 2014 & Year <= 2024)

# Write 'last_10_years' dataset to a CSV file
write.csv(last_10_years, file = "last_10_years.csv", row.names = FALSE)

# Create bar graph
barplot(last_10_years$Temperature, 
        names.arg = last_10_years$Year,
        xlab = "Year",
        ylab = "Average Temperature (Fahrenheit)",
        main = "Average Annual Temperature in Pennsylvania (2014-2024)",
        ylim = c(40, 80), # Set the y-axis limits to match the range of temperatures
        col = "skyblue", # Set bar color
        border = NA, # Remove border of bars
        space = 0.5, # Adjust spacing between bars
        cex.names = 0.6) # Adjust font size of year labels

```
> Data within this figure is randomized and **not accurate** to actual annual temperatures in Pennsylvania, practice making bar charts

##The following graphs are using previous years of peatland data

```{r}

# Load libraries
library(ggplot2)
library(dplyr)

# Create data frame for Carbon concentration in 2018
Carbon_2018_data <- data.frame(
  month = c("6", "7", "8"),
  Concentration = c(444.7, 531.9, 450.6)
)

# Convert 'month' column to numeric
Carbon_2018_data$month <- as.numeric(Carbon_2018_data$month)

# Plot a line graph
ggplot(data = Carbon_2018_data, aes(x = month, y = Concentration)) +
  geom_line(color = "#1F77B4", size = 1) +  # Add a line
  geom_point(color = "#1F77B4", size = 3) +  # Add points
  ggtitle("Carbon (C) Concentration in Everiam Plant Tissue in the Summer of 2018") +
  ylab("C Concentration (mg/g)") +
  xlab("Month") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(color = "black"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  )

```

>Chart displays concentrations of C found in everiam plant tissue at one site (Kearl) over 3 months in the summer of 2018.

##Average C Concentration in 2018 and 2019; Everiam Plant Tissue, Kearl site
```{r}
avg_c_year <- data.frame(
  average = c(442.4, 436),
  year = c("2018", "2019")
)

# Load libraries
library(ggplot2)

# Create bar graph with modified aesthetic
ggplot(data = avg_c_year, aes(x = year, y = average, fill = year)) + 
  geom_bar(stat = "identity", color = "white", width = 0.5) +  # Adjust width and remove outline
  ggtitle("Average Carbon (C) Concentration in Everiam Plant Tissue Across Five Plots") +
  ylab("C Concentration (mg/g)") +
  xlab("Year") +
  scale_fill_manual(values = c("#1F77B4", "#1F77B4")) +  # Set bars to the same blue color
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(color = "black"),
    legend.position = "none",
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  )


```

> Comparing average concentrations of carbon found in everiam plant tissue at the same site (Kearl) in 2018 versus 2019.

```{r}

# Load libraries
library(ggplot2)
library(tidyr)

# Create the data frame with concentrations
avg_CNS_year <- data.frame(
  C = c(442.3, 436.0),
  N = c(8.37, 7.2),
  S = c(1.5, 1.9),
  year = c("2018", "2019")
)

# Reshape data from wide to long format
avg_CNS_year_long <- pivot_longer(avg_CNS_year, cols = c(C, N, S), names_to = "Element", values_to = "Value")

# Create bar graph with modern aesthetic
ggplot(avg_CNS_year_long, aes(x = year, y = Value, fill = Element)) +
  geom_bar(stat = "identity", position = "dodge", color = "white", width = 0.7) +
  labs(title = "Average Concentrations of C, N, and S Over Years",
       x = "Year",
       y = "Concentration (mg/g)") +
  scale_fill_manual(values = c("C" = "#1F77B4", "N" = "#2CA02C", "S" = "#FF7F0E"),
                    name = "Element") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(color = "black"),
    legend.position = "top",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  )


```
> Average concentrations of C, N, and S in Everiam plant tissue in Kearl Canada in 2018 and 2019

```{r}

# Load libraries
library(ggplot2)

# Create a custom color palette
my_colors <- c("N" = "#2CA02C", "S" = "#FF7F0E")

# Create bar graph with modern aesthetic
ggplot(avg_CNS_year_long[avg_CNS_year_long$Element %in% c("N", "S"), ], 
       aes(x = year, y = Value, fill = Element)) +
  geom_bar(stat = "identity", position = "dodge", color = "white", width = 0.7) +
  ggtitle("Average Concentrations of N and S Over Years") +
  ylab("Concentration") +
  xlab("Year") +
  scale_fill_manual(values = my_colors, name = "Element") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(color = "black"),
    legend.position = "top",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  )

```
> A closer look at N and S concentration in Everiam plant tissue in 2018 and 2019

```{r}
# Load libraries
library(ggplot2)
library(dplyr)

# Create data frame for N and S concentrations
N_S_data <- data.frame(
  month = c("6", "7", "8"),
  N_Concentration = c(8.37, 7.2, 6.5),
  S_Concentration = c(1.5, 1.9, 2.2)
)

# Convert 'month' column to numeric
N_S_data$month <- as.numeric(N_S_data$month)

# Plot a line graph with a similar aesthetic
ggplot(data = N_S_data, aes(x = month)) +
  geom_line(aes(y = N_Concentration, color = "N"), size = 1) +
  geom_line(aes(y = S_Concentration, color = "S"), size = 1) +
  ggtitle("Average Concentrations of N and S over summer of 2018") +
  ylab("Concentration (mg/g)") +
  xlab("Month") +
  scale_color_manual(name = "Element", values = c(N = "#2CA02C", S = "#FF7F0E")) +  # Set custom colors
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(color = "black"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  )

```
> Average concentrations of N and S found in everiam plant tissue at one site (Kearl) pver the summer of 2018.
